---
title: "Capstone"
author: "Abhishek & Sajjiynie"
date: "2024-03-28"
output:
  html_document:
    css: styles.css
    code_folding: show
    anchor_sections: true
    toc: true

---
```{css, echo=FALSE}
body {
  background-color: #10101010; 
  color: #333333; 
}
.fold-show::before {
  content: 'Show Code';
}
.fold-hide::before {
  content: 'Hide Code';
}

```



# Loading All necessary Libraries


```{r}

library(DataExplorer)
library(plumber)
library(readr)
library(tidyverse)
library(caret)
library(randomForest)
library(glmnet)
library(ggplot2)
library(dplyr)
library(scales)
library(cowplot)
library(corrplot)
library(MatchIt)
library(caTools)
library(glm2)
library(reshape2)
library(pROC)
library(car)
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(plotly)
library(cluster)



ab_dataset_cap <- read_csv("ibmdatacleaned.csv")

```

# Dataset Exploration

```{r echo = TRUE, results = 'hide'}

data.frame(ab_dataset_cap)

```


# Variable conversion for Numeric and Categorical variables

```{r}

# Conversion of Numeric Variables
numeric_vars <- c("Age", "DailyRate", "DistanceFromHome", "Education", "EmployeeNumber", 
                  "EnvironmentSatisfaction", "HourlyRate", "JobInvolvement", "JobLevel", 
                  "JobSatisfaction", "MonthlyIncome", "MonthlyRate", "NumCompaniesWorked", 
                  "PercentSalaryHike", "PerformanceRating", "RelationshipSatisfaction", 
                  "StandardHours", "StockOptionLevel", "TotalWorkingYears", 
                  "TrainingTimesLastYear", "WorkLifeBalance", "YearsAtCompany", 
                  "YearsInCurrentRole", "YearsSinceLastPromotion", "YearsWithCurrManager", "Index")

# Ensure that the variables actually exists in the dataframe before conversion
existing_numeric_vars <- numeric_vars[numeric_vars %in% names(ab_dataset_cap)]

# Convert existing variables to numeric
ab_dataset_cap[existing_numeric_vars] <- lapply(ab_dataset_cap[existing_numeric_vars], function(x) as.numeric(as.character(x)))

# Update: Convert categorical character variables to factors (As previously defined)
categorical_vars <- c("Department", "EducationField", "JobRole", "Attrition", "BusinessTravel",
                      "Education Desc", "EnvironmentSatisfaction Desc", "Gender", 
                      "JobInvolvement Desc", "JobSatisfaction Desc", "MaritalStatus", 
                      "OverTime", "WorkLifeBalance Desc", "Employee Source")

# Filter existing categorical variables
existing_categorical_vars <- categorical_vars[categorical_vars %in% names(ab_dataset_cap)]

ab_dataset_cap$`RelationshipSatisfaction Desc` <- as.factor(ab_dataset_cap$`RelationshipSatisfaction Desc`)


# Convert to factor
ab_dataset_cap[existing_categorical_vars] <- lapply(ab_dataset_cap[existing_categorical_vars], factor)

# Print structure to verify changes
str(ab_dataset_cap)

```


# Missing Values 


```{r}
# Calculate the number of missing values in each column
missing_values <- sapply(ab_dataset_cap, function(x) sum(is.na(x)))
missing_values
```


#Summary of Missing Values

```{r corrected-summary-missing-values}


# Summary of missing (NA) values
na_summary <- ab_dataset_cap %>%
  summarise_all(~sum(is.na(.))) %>%
  mutate(Source = "NA")

# If interested in zeros as potential placeholders for missing data
zero_summary <- ab_dataset_cap %>%
  summarise_all(~sum(. == 0, na.rm = TRUE)) %>%
  mutate(Source = "Zero")

# Combine summaries
summary_combined <- bind_rows(na_summary, zero_summary)

print(summary_combined)


```



# Attrition Class and Distribution


```{r}

# Calculate total count for normalization
total_count_numeric <- sum(table(ab_dataset_cap$Attrition))

# Visualize the distribution of Attrition_numeric with numbers and rounded percentages
ggplot(ab_dataset_cap, aes(x = factor(Attrition), fill = factor(Attrition))) +
  geom_bar() +
  geom_text(
    aes(label = paste(..count.., " (", round(..count../total_count_numeric*100, 2), "%)", sep="")),
    stat = 'count', 
    position = position_stack(vjust = 0.5)
  ) +
  theme_minimal() +
  labs(title = "Distribution of Attrition", x = "Attrition Status", y = "Count") +
  scale_y_continuous(labels = scales::comma)



```


# Column Names


```{r}
colnames(ab_dataset_cap)
```


# Data Profiling Report


```{r eda-manual-sections-improved, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=10}


# Set plot size larger for Jupyter notebooks/RStudio Viewer to prevent label overlap
options(repr.plot.width = 12, repr.plot.height = 10)

# Dataset introduction
introduce(ab_dataset_cap)

# Data structure
str(ab_dataset_cap)

# Missing value profile
plot_missing(ab_dataset_cap)

# Univariate distribution for all features

selected_vars <- names(ab_dataset_cap)[1:10]  # Adjust the index to select variables
plot_histogram(ab_dataset_cap[, selected_vars])

# Correlation analysis

plot_correlation(ab_dataset_cap[, selected_vars])

# Principal Component Analysis for dimensionality reduction visualization

plot_prcomp(ab_dataset_cap, maxcat = 5L)

# Bivariate distribution with target variable (if Attrition is a factor)
plot_boxplot(ab_dataset_cap, by = "Attrition")

```


# Conversion of Variables for modelling


```{r}

# Convert Attrition to a binary numeric variable
ab_dataset_cap$Attrition_numeric <- as.numeric(ab_dataset_cap$Attrition == "Voluntary Resignation")

# Calculate correlations with Attrition
correlations <- cor(ab_dataset_cap[, sapply(ab_dataset_cap, is.numeric)])
cor_attrition <- correlations[,"Attrition_numeric"]
sorted_cor <- sort(cor_attrition, decreasing = TRUE)

# View the correlations sorted
print(sorted_cor)

```

# Interpretations of initial correlation matrix

DistanceFromHome (0.069913505):
Positive correlation (0.07) with Attrition_numeric suggests that there is a slight tendency for employees who live farther from the workplace to have a slightly higher likelihood of attrition.

NumCompaniesWorked (0.040475987):
Positive correlation (0.04) with Attrition_numeric suggests that employees who have worked at more companies before joining the current one may have a slightly higher tendency to leave.

HourlyRate (0.012403772):
Very weak positive correlation (0.01) with Attrition_numeric, indicating that there's almost no relationship between hourly rate and attrition.

Index (-0.002001987):
This variable seems to have almost no correlation with Attrition_numeric or other variables, indicated by a very close to zero coefficient.

EmployeeNumber (-0.002019309):
Almost no correlation with Attrition_numeric or other variables.

RelationshipSatisfaction (-0.005611355):
Very weak negative correlation (-0.01) with Attrition_numeric, suggesting that employees with higher relationship satisfaction might have a slightly lower likelihood of attrition.

PerformanceRating (-0.006735761):
Very weak negative correlation (-0.01) with Attrition_numeric, implying that employees with higher performance ratings might have a slightly lower likelihood of attrition.

MonthlyRate (-0.008343636):
Very weak negative correlation (-0.01) with Attrition_numeric, indicating that there's almost no relationship between monthly rate and attrition.

YearsSinceLastPromotion (-0.018987128):
Weak negative correlation (-0.02) with Attrition_numeric, suggesting that employees who have been promoted more recently may have a slightly lower likelihood of attrition.




# Initial variable selection

```{r create-subset}
# Creating a subset by excluding specified columns
ab_dataset_reduced <- ab_dataset_cap[, !(names(ab_dataset_cap) %in% c(
  "Education Desc", 
  "EnvironmentSatisfaction Desc", 
  "JobInvolvement Desc", 
  "JobSatisfaction Desc", 
  "RelationshipSatisfaction Desc", 
  "WorkLifeBalance Desc",
  "EmployeeNumber",
  "Unique Emp ID",
  "Application ID",
  "Index",
  "StandardHours",
  "Attrition",
  "Employee Source"
))]

```


# Correlation Matrix heatmap



```{r new - correlation}



# Calculate correlation matrix for numeric variables
numeric_vars <- sapply(ab_dataset_reduced, is.numeric)
cor_matrix <- cor(ab_dataset_reduced[, numeric_vars], use = "pairwise.complete.obs")

# Melt the correlation matrix into a long format
melted_cor_matrix <- melt(cor_matrix)

# Create the heatmap with ggplot2
ggplot(data = melted_cor_matrix, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile() +
  scale_fill_gradient2(name = "Correlation", low = "blue", mid = "white", high = "red", midpoint = 0) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        axis.text.y = element_text(angle = 0),
        axis.title.x = element_blank(),
        axis.title.y = element_blank()) +
  labs(title = "Correlation Matrix Heatmap with initial selected Variables")



```

# Re checking the variable names in reduced dataset


```{r}

colnames(ab_dataset_reduced)

```

# Attrition Class Distribution


```{r}


# Create a more appealing bar plot using ggplot2
ggplot(ab_dataset_reduced, aes(x=factor(Attrition_numeric), fill=factor(Attrition_numeric))) + 
  geom_bar(stat="count", position="dodge", color="black") +
  geom_text(stat='count', aes(label=..count..), vjust=-0.5, color="black") +
  scale_fill_manual(values=c("#56B4E9", "#E69F00"), labels=c("No Attrition (0)", "Attrition (1)")) +
  scale_y_continuous(labels=scales::comma) +
  labs(title="Attrition Numeric Distribution", x="Attrition Status", y="Frequency", fill="Attrition Status") +
  theme_minimal() +
  theme(text = element_text(size=12, face="bold"), 
        plot.title = element_text(hjust = 0.5), 
        legend.title = element_blank(), 
        axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid.major = element_line(size = 0.1, linetype = 'solid', color = "gray"),
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "white", color = "white"),
        legend.position = "bottom")

```


# Data Pre-processing Steps


```{r}

# Convert 'Attrition_numeric' to a factor
ab_dataset_reduced$Attrition_numeric <- as.factor(ab_dataset_reduced$Attrition_numeric)

# Check for multicollinearity
cor_matrix <- cor(ab_dataset_reduced[, sapply(ab_dataset_reduced, is.numeric)], use = "pairwise.complete.obs")

# Identify high correlations excluding self-correlations
high_cor <- which(abs(cor_matrix) > 0.7 & cor_matrix != 1, arr.ind = TRUE)

high_cor_vars <- data.frame(
  Var1 = rownames(cor_matrix)[high_cor[, 1]],
  Var2 = colnames(cor_matrix)[high_cor[, 2]],
  Correlation = cor_matrix[high_cor]
)

print(high_cor_vars)
```



# Plotting positive correlation 1


```{r}

# Positive Correlations
p1 <- ab_dataset_reduced %>% select(TotalWorkingYears, MonthlyIncome) %>%
ggplot(aes(x=TotalWorkingYears, y=MonthlyIncome)) + 
  geom_point(colour = "#6D98BA", alpha=0.5) + 
  geom_smooth(method="loess", color="#D95F02") + 
  theme_minimal() + 
  theme(legend.position="bottom", 
        plot.title=element_text(hjust=0.5), 
        plot.subtitle=element_text(hjust=0.5)) + 
  labs(title="Positive Correlation", subtitle="Monthly Income vs Working Years")



# Arrange the plots into a grid
plot_grid(p1)

```


# Attrition Numeric against years in current role



```{r}
# YearsInCurrentRole vs Attrition_numeric
mean_years_in_current_role <- mean(ab_dataset_reduced$YearsInCurrentRole)
p2 <- ab_dataset_reduced %>% select(YearsInCurrentRole, Attrition_numeric) %>%
  ggplot(aes(x = YearsInCurrentRole, fill = Attrition_numeric)) +
  geom_histogram(position = "dodge", bins = 20, color = "black", alpha = 0.7) +
  geom_vline(xintercept = mean_years_in_current_role, linetype = "dashed", size = 1, color = "red", 
             aes(label = paste("Mean:", round(mean_years_in_current_role, 2)))) +
  scale_color_manual(values = c("red")) +
  theme_minimal() +
  labs(title = "Attrition by Years in Current Role", x = "Years in Current Role", y = "Count", fill = "Attrition") +
  theme(legend.position = "top")
print(p2)

```


# Training times Last year vs Attrition numeric


```{r}
# TrainingTimesLastYear vs Attrition_numeric
mean_training_times_last_year <- mean(ab_dataset_reduced$TrainingTimesLastYear)
p3 <- ab_dataset_reduced %>% select(TrainingTimesLastYear, Attrition_numeric) %>%
  ggplot(aes(x = TrainingTimesLastYear, fill = Attrition_numeric)) +
  geom_histogram(position = "dodge", bins = 20, color = "black", alpha = 0.7) +
  geom_vline(xintercept = mean_training_times_last_year, linetype = "dashed", size = 1, color = "red", 
             aes(label = paste("Mean:", round(mean_training_times_last_year, 2)))) +
  scale_color_manual(values = c("red")) +
  theme_minimal() +
  labs(title = "Attrition by Training Times Last Year", x = "Training Times Last Year", y = "Count", fill = "Attrition") +
  theme(legend.position = "top")
print(p3)

```

# Distance from home vs Attrition numeric

```{r}
# DistanceFromHome vs Attrition_numeric
mean_distance_from_home <- mean(ab_dataset_reduced$DistanceFromHome)
p4 <- ab_dataset_reduced %>% select(DistanceFromHome, Attrition_numeric) %>%
  ggplot(aes(x = DistanceFromHome, fill = Attrition_numeric)) +
  geom_histogram(position = "dodge", bins = 20, color = "black", alpha = 0.7) +
  geom_vline(xintercept = mean_distance_from_home, linetype = "dashed", size = 1, color = "red", 
             aes(label = paste("Mean:", round(mean_distance_from_home, 2)))) +
  scale_color_manual(values = c("red")) +
  theme_minimal() +
  labs(title = "Attrition by Distance from Home", x = "Distance from Home", y = "Count", fill = "Attrition") +
  theme(legend.position = "top")
print(p4)

```

# Calculating Mean age across different departments 

```{r}
mean_ages <- ab_dataset_reduced %>%
  group_by(Attrition_numeric) %>%
  summarise(
    MeanAge = mean(Age, na.rm = TRUE),
    MaxCount = max(table(Age))
  )

```




# Calculate the mean age and the maximum count for each attrition status

```{r positive_correlation_totalworkingyears, fig.cap="Total Working Years vs Attrition"}


# Compute mean ages and max count for each attrition status
mean_ages <- ab_dataset_reduced %>%
  group_by(Attrition_numeric) %>%
  summarise(
    MeanAge = mean(Age, na.rm = TRUE),
    MaxCount = max(table(Age))
  )

# Create the histogram
age_distribution_plot <- ab_dataset_reduced %>%
  ggplot(aes(x = Age, fill = factor(Attrition_numeric))) + 
  geom_histogram(binwidth = 1, position = 'identity', alpha = 0.6) +
  scale_fill_manual(values = c("#5DA5B3", "#26547C"), 
                    labels = c("Stayed", "Left"),
                    name = "Attrition Status") +
  geom_vline(data = mean_ages,
             aes(xintercept = MeanAge, color = factor(Attrition_numeric)),
             linetype = "dashed",
             size = 1.5) +
  geom_text(data = mean_ages, 
            aes(x = MeanAge, label = paste("Mean:", round(MeanAge, 1)), y = MaxCount, color = factor(Attrition_numeric)),
            size = 4, vjust = 1.5, fontface = "bold", color = "black") +
  scale_color_manual(values = c("#5DA5B3", "#26547C")) +
  guides(color = FALSE) +
  theme_minimal() +
  labs(title = "Age Distribution by Attrition Status",
       x = "Age",
       y = "Count") +
  theme(legend.position = "top",
        plot.title = element_text(size = 14, face = "bold"),
        axis.title = element_text(size = 12),
        legend.background = element_blank(),
        legend.key = element_blank())



# Print the interactive plot
print(age_distribution_plot)


```


# Age vs Monthly Income


```{r}

# Positive Correlations
p1 <- ab_dataset_reduced %>% select(Age, MonthlyIncome) %>%
  ggplot(aes(x=Age, y=MonthlyIncome)) + 
  geom_point(colour = "#6D98BA", alpha=0.5) + 
  geom_smooth(method="loess", color="#D95F02") + 
  theme_minimal() + 
  theme(legend.position="bottom", 
        plot.title=element_text(hjust=0.5), 
        plot.subtitle=element_text(hjust=0.5)) + 
  labs(title="Positive Correlation", subtitle="Monthly Income vs Age")

# Arrange the plots into a grid
plot_grid(p1)

# Define the color palette
color_palette <- brewer.pal(9, "Blues")

```




```{r}

mean_ages_dept <- ab_dataset_reduced %>%
  group_by(Department, Attrition_numeric) %>%
  summarise(
    MeanAge = mean(Age, na.rm = TRUE),
    MaxCount = max(table(Age)),
    .groups = 'drop' # This will drop the grouping so that we don't carry it into our plot
  )


```



# Age Distribution


```{r}
age_distribution_plot_dept <- ab_dataset_reduced %>%
  ggplot(aes(x = Department, fill = factor(Attrition_numeric))) + 
  geom_bar(position = 'dodge', alpha = 0.6) +
  scale_fill_manual(values = c("#6D88BA", "#D93F02"), 
                    labels = c("Stayed", "Left"),
                    name = "Attrition Status") +
  geom_text(data = mean_ages_dept, 
            aes(x = Department, y = MaxCount, 
                label = paste("Mean:", round(MeanAge, 1)), 
                color = factor(Attrition_numeric)),
            size = 4, vjust = -0.5, fontface = "bold", 
            position = position_dodge(width = 0.9)) +
  scale_color_manual(values = c("black", "black")) + # Set both colors to white
  theme_minimal() +
  labs(title = "Attrition Distribution by Department",
       x = "Department",
       y = "Count") +
  theme(legend.position = "top",
        plot.title = element_text(size = 14, face = "bold"),
        axis.title = element_text(size = 12),
        legend.text = element_text(color = "black"), # Ensure legend text is black for visibility
        legend.background = element_blank(),
        legend.key = element_blank(),
        legend.title = element_text(color = "black")) + # Ensure legend title is black for visibility
  guides(color = FALSE) # Remove the color guide for the text

# Print the plot
print(age_distribution_plot_dept)


```




# Scaling the numerical Variables


```{r}

# Scale the numeric variables in the dataset "ab_dataset_reduced"
numeric_vars <- sapply(ab_dataset_reduced, is.numeric)
ab_dataset_reduced[numeric_vars] <- scale(ab_dataset_reduced[numeric_vars])

# Ensure the dataset is in a data frame format
ab_dataset_reduced <- as.data.frame(ab_dataset_reduced)

# View the structure of the dataset to confirm changes
str(ab_dataset_reduced)


```

# Summary of numerical Variables

```{r}
summary(ab_dataset_reduced[numeric_vars])
```



# Splitting the Dataset into Training and Testing

```{r}

set.seed(123)  # for reproducible random sampling

# Split the data into training and testing sets
index <- createDataPartition(y = ab_dataset_reduced$Attrition_numeric, p = 0.7, list = FALSE)
training_set <- ab_dataset_reduced[index, ]
testing_set <- ab_dataset_reduced[-index, ]

# Check the number of cases in each set
nrow(training_set)  # Should be approximately 70% of the data
nrow(testing_set)   # Should be approximately 30% of the data

# Convert Attrition_numeric to a factor with explicit levels in both the training and testing sets
training_set$Attrition_numeric <- factor(training_set$Attrition_numeric, levels = c(0, 1))
testing_set$Attrition_numeric <- factor(testing_set$Attrition_numeric, levels = c(0, 1))


```



# Balancing the Attrition_numeric Class

```{r}


# Check initial class distribution
table(training_set$Attrition_numeric)


# Balance the classes
training_set <- downSample(x = training_set[, -which(names(training_set) == "Attrition_numeric")],
                                    y = training_set$Attrition_numeric)

# Check the new class distribution to confirm balancing
table(training_set$Class)

# Renaming 'Class' back to 'Attrition_numeric'
training_set$Attrition_numeric <- training_set$Class
training_set$Class <- NULL 


```




# Visual Representation of Attrition after balancing

```{r}

# Calculate counts and percentages
attrition_counts <- table(training_set$Attrition_numeric)
attrition_df <- data.frame(Attrition_numeric = names(attrition_counts), 
                           Count = as.integer(attrition_counts))
attrition_df$Percentage <- (attrition_df$Count / sum(attrition_df$Count)) * 100

# Create the bar plot
ggplot(attrition_df, aes(x = Attrition_numeric, y = Count, fill = Attrition_numeric)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = paste(Count, " (", sprintf("%.1f", Percentage), "%)", sep="")), 
            vjust = -0.5, position = position_stack(vjust = 0.5)) +
  labs(title = "Distribution of Attrition Numeric", x = "Attrition Numeric Status", y = "Count") +
  theme_minimal()

```



```{r}
names(training_set)
```




# Fitting Logistic model 1 on training set 


```{r}


# Fit logistic regression model on training set
logistic_model <- glm(Attrition_numeric ~ ., data = training_set, family = binomial())

# Print out the results of the logistic regression model
summary(logistic_model)

```

# Logistic model 1 on test set & confusion matrix


```{r}

# Predicting on the testing set
testing_set_predictions <- predict(logistic_model, newdata = testing_set, type = "response")

# Converting probabilities to binary class predictions based on a threshold (usually 0.5)
predicted_classes <- ifelse(testing_set_predictions > 0.5, 1, 0)

# Create a confusion matrix
conf_mat <- confusionMatrix(as.factor(predicted_classes), as.factor(testing_set$Attrition_numeric))

# Print the confusion matrix along with statistics
print(conf_mat)


```


# Interpretation for Logistic model 1

Reference:
0 represents the negative class (e.g., employees who do not leave the company).
1 represents the positive class (e.g., employees who leave the company).

Prediction:
0 represents the model's prediction of the negative class.
1 represents the model's prediction of the positive class.

True Positives (TP): 3888
The model correctly predicted that 3888 instances belong to the positive class (employees who leave) and indeed they belong to the positive class.

False Positives (FP): 321
The model incorrectly predicted that 321 instances belong to the positive class when they actually belong to the negative class (Type I error).

True Negatives (TN): 776
The model correctly predicted that 776 instances belong to the negative class (employees who don't leave) and indeed they belong to the negative class.

False Negatives (FN): 1961
The model incorrectly predicted that 1961 instances belong to the negative class when they actually belong to the positive class (Type II error).

Accuracy: 0.6715

Overall proportion of correct predictions made by the model.

Precision (Positive Predictive Value): 0.9237
Proportion of instances predicted as positive by the model that are actually positive.

Recall (Sensitivity): 0.6647
Proportion of actual positive instances that were correctly predicted by the model.

Specificity: 0.7074
Proportion of actual negative instances that were correctly predicted by the model.

F1 Score:
Harmonic mean of precision and recall, providing a balance between the two metrics.

Kappa: 0.2315
A statistic that measures inter-rater agreement for categorical items.

Detection Rate: 0.5597
Rate of correct predictions among all actual positive instances.

Balanced Accuracy: 0.6861
Average of sensitivity and specificity, useful for imbalanced datasets.




# AUC for Logistic Model


```{r}

# Calculate AUC
roc_result <- roc(response = testing_set$Attrition_numeric, predictor = as.numeric(testing_set_predictions))
auc_value <- auc(roc_result)

# Print AUC
print(auc_value)

```


# AUC plot


```{r}


# Calculate ROC curve
roc_result <- roc(response = testing_set$Attrition_numeric, predictor = as.numeric(testing_set_predictions))

# Calculate AUC
auc_value <- auc(roc_result)

# Create a data frame from the roc object
roc_data <- data.frame(
  TPR = roc_result$sensitivities,    # True Positive Rate (Sensitivity)
  FPR = roc_result$specificities,    # False Positive Rate (1 - Specificity)
  thresholds = roc_result$thresholds
)


# Plot ROC curve using ggplot2
ggroc_plot <- ggplot(data = roc_data, aes(x = 1 - FPR, y = TPR)) +
  geom_line(color = "blue", size = 1.5) +
  geom_ribbon(aes(ymin = 0, ymax = TPR), fill = "blue", alpha = 0.2) +  # Adds shaded area under curve
  geom_abline(linetype = "dashed", color = "red") +  # Adds the reference line
  ggtitle("ROC Curve with AUC") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),  # Centers the title
        axis.title = element_text(size = 12),    # Adjust the axis titles size
        axis.text = element_text(size = 12)) +   # Adjust the axis text size
  xlab("False Positive Rate (1 - Specificity)") +
  ylab("True Positive Rate (Sensitivity)") +
  annotate("text", x = .5, y = .2, label = paste("AUC =", round(auc_value, 2)), 
           size = 5, color = "black", hjust = 0.5)

# Print the plot
print(ggroc_plot)

```


# Correlation plot after logistic model


```{r}

numeric_data <- training_set[, sapply(training_set, is.numeric)]

# Calculate the correlation matrix
cor_matrix <- cor(numeric_data)

# Identify highly correlated pairs
highly_correlated_pairs <- findCorrelation(cor_matrix, cutoff = 0.7, verbose = TRUE)

# Print out highly correlated pairs
print(highly_correlated_pairs)

# Visualize the correlation matrix
corrplot(cor_matrix, method = "circle", type = "upper", order = "hclust",
         tl.col = "black", tl.srt = 45, tl.cex = 0.6, cl.cex = 0.6)

```



# LASSO Regression for Variable significance


```{r}

# Create a matrix of predictors and a response variable vector
x <- model.matrix(Attrition_numeric ~ . - 1, data = training_set)  # -1 to exclude intercept
y <- training_set$Attrition_numeric


```


```{r}
set.seed(123)  # For reproducible results
cv_model <- cv.glmnet(x, y, alpha = 1, family = "binomial")  # alpha = 1 for LASSO

```



```{r}

best_lambda <- cv_model$lambda.min

```


```{r}

final_model <- glmnet(x, y, alpha = 1, lambda = best_lambda, family = "binomial")

```



```{r}
# Print non-zero coefficients
print(coef(final_model))

```


# Interpretation of coefficients (lasso)
 
(Intercept): The model intercept or bias term is approximately -1.666, which is the log-odds of the outcome when all predictor variables are zero.

Coefficients Close to Zero (.): These are variables for which LASSO has effectively shrunk the coefficient to zero, suggesting that they do not contribute significantly to the model. The dot (.) is a placeholder indicating no effect after the penalty has been applied.

Positive Coefficients: Variables such as DepartmentSales (0.166) and BusinessTravelTravel_Frequently (1.336) have positive coefficients, indicating that higher values of these variables are associated with an increased log-odds of the outcome (possibly higher attrition, if Attrition_numeric = 1 represents attrition). For instance, employees in the sales department or those who travel frequently are more likely to experience the event of interest (which might be attrition in this context).

Negative Coefficients: Variables like JobRoleHuman Resources Manager (-3.278) have negative coefficients, indicating that higher values of these variables are associated with a decreased log-odds of the outcome. This might suggest that being in a human resources manager role is associated with a lower likelihood of attrition.

Large Positive Coefficients: Some coefficients like OverTimeYes (0.865) are substantially positive, which would suggest a strong association with the outcome. Employees who work overtime might have a much higher likelihood of attrition.
Significant Predictors After Shrinkage: The LASSO method penalizes the absolute size of the coefficients and can drive some to zero, effectively selecting variables by removing those with negligible effects. The remaining non-zero coefficients are the variables LASSO has selected as having a more significant relationship with the outcome.



# Removing variables that do not contribute significantly




```{r}

# Define all variables to be removed in one list
variables_to_remove <- c("TotalWorkingYears", "YearsWithCurrManager", "YearsInCurrentRole",
  "MonthlyRate", "HourlyRate", "DailyRate", "JobLevel", "JobRole")

# Removing the specified variables from the dataset
ab_dataset_reduced <- ab_dataset_reduced[, !(names(ab_dataset_reduced) %in% variables_to_remove)]

# Confirm the removal by printing the remaining variable names
print(names(ab_dataset_reduced))


```




# Logistic Model 2 - with selected variables


```{r}

set.seed(123) # for reproducibility

# Create a partition to split the dataset into training and testing sets
split <- createDataPartition(ab_dataset_reduced$Attrition_numeric, p = 0.7, list = FALSE)

# Create the training and testing datasets
trainingdata_reduced_ab <- ab_dataset_reduced[split, ]
testingdata_reduced_ab <- ab_dataset_reduced[-split, ]



# Balance the Attrition_numeric class in the training set
set.seed(123) # for reproducibility

trainingdata_balanced_ab <- upSample(x = trainingdata_reduced_ab[, -which(names(trainingdata_reduced_ab) == "Attrition_numeric")],
                                     y = trainingdata_reduced_ab$Attrition_numeric)

# Check the balance of the updated training set
table(trainingdata_balanced_ab$Class)


```


# Fitting the logistic model 2


```{r}

# Fit the logistic regression model on the balanced training set
reduced_logistic_model_ab <- glm(Class ~ ., data = trainingdata_balanced_ab, family = binomial())

# Rename 'Class' back to 'Attrition_numeric'
names(trainingdata_balanced_ab)[names(trainingdata_balanced_ab) == "Class"] <- "Attrition_numeric"

# Fit the logistic regression model using 'Attrition_numeric' as the dependent variable
reduced_logistic_model_ab_updated <- glm(Attrition_numeric ~ ., data = trainingdata_balanced_ab, family = binomial())

# Print out the summary of the updated logistic regression model
summary(reduced_logistic_model_ab_updated)


```

```{r}
# Calculate Odds Ratios
exp(coef(reduced_logistic_model_ab_updated))
```

# Odds ratio interpretation

(Intercept): The odds ratio for the intercept is 0.1637746. Since the intercept represents the log odds of the outcome when all predictors are held at zero, this value isn't often used for interpretation in the context of the predictors.

DepartmentResearch & Development: The odds of the outcome (likely an event such as employee attrition, based on the context of the variable names) for employees in the Research & Development department are 0.9872168 times those of the baseline department (not shown here, but typically the one excluded from the model), holding all other variables constant. This value is very close to 1, indicating almost no change in odds compared to the baseline.

DepartmentSales: Employees in the Sales department have 1.4792136 times the odds of the outcome compared to the baseline department, holding all other variables constant. This suggests a 47.9% increase in the odds of the outcome occurring.
Age: For each one-unit increase in age, the odds of the outcome decrease by a factor of 0.6736967 (or 32.6% decrease in the odds), holding other variables constant.

BusinessTravelTravel_Frequently: Employees who travel frequently have 3.9708247 times the odds of the outcome compared to those who do not travel as part of their job (likely the reference category), all else being equal. This is a substantial increase in odds.

MonthlyIncome: For each unit increase in monthly income (the scale isn’t specified here, so it could be dollars, thousands of dollars, etc.), the odds of the outcome are multiplied by 0.8816012, indicating a decrease in the odds of the outcome as monthly income increases.

OverTimeYes: Employees who work overtime have 2.4904832 times the odds of the outcome happening compared to those who do not work overtime, controlling for other factors.

MaritalStatusSingle: Employees who are single have 1.7334793 times the odds of the outcome compared to the baseline marital status (likely married or divorced, which is not listed).



```{r}

# Make predictions on the testing set using the logistic regression model
testingdata_predictions_ab <- predict(reduced_logistic_model_ab_updated, newdata = testingdata_reduced_ab, type = "response")

# Convert predictions to a binary outcome based on a threshold (e.g., 0.5)
predicted_class_ab <- ifelse(testingdata_predictions_ab > 0.5, 1, 0)

# Actual values of Attrition_numeric from the testing set
actual_class_ab <- testingdata_reduced_ab$Attrition_numeric

# Confusion Matrix to see how well the model did on the testing set
confusionMatrix(data = as.factor(predicted_class_ab), reference = as.factor(actual_class_ab))

```



# ROC Curve for Logistic model 2


```{r refined-roc-curve, echo=TRUE, message=FALSE, warning=FALSE}

# Generate the ROC object from actual and predicted classes
roc_object <- roc(response = actual_class_ab, predictor = testingdata_predictions_ab)

# Create a data frame from the roc_object for ggplot
roc_data <- data.frame(
  True_Positive_Rate = roc_object$sensitivities,
  False_Positive_Rate = roc_object$specificities,
  Threshold = roc_object$thresholds
)

# Calculate AUC
auc_value <- auc(roc_object)

# ROC Curve plot with ggplot2
roc_curve_plot <- ggplot(roc_data, aes(x = 1 - False_Positive_Rate, y = True_Positive_Rate)) +
  geom_line(color = "#1c61b6", size = 1) +
  geom_segment(aes(x = 0, y = 0, xend = 1, yend = 1), color = "red", linetype = "dotted") +
  geom_text(aes(x = 0.8, y = 0.2, label = paste("AUC =", round(auc_value, 3))), hjust = 0, vjust = 0, color = "red") +
  labs(title = "ROC Curve with AUC", x = "False Positive Rate (1 - Specificity)", y = "True Positive Rate (Sensitivity)") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        text = element_text(size = 12, face = "bold"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "white", color = "white"),
        legend.position = "none")

# Print the ROC Curve plot
print(roc_curve_plot)

```








# Department wise (Human Resources, Sales and Research and Development) Logistic Regressions
 
 
```{r}

# Split the subset data into three separate data frames, one for each department
hr_subset <- subset(ab_dataset_reduced, Department == "Human Resources")
sales_subset <- subset(ab_dataset_reduced, Department == "Sales")
rd_subset <- subset(ab_dataset_reduced, Department == "Research & Development")

# Function to check factor levels in each subset, revised to exclude non-factors and return only factors
check_factor_levels <- function(data) {
  lapply(data[, sapply(data, is.factor)], function(x) length(levels(x)))
}

# Apply to each department
hr_levels <- check_factor_levels(hr_subset)
sales_levels <- check_factor_levels(sales_subset)
rd_levels <- check_factor_levels(rd_subset)


```



# Sales Subset


```{r}

# Subset data for the Sales department from main dataset
sales_subset <- subset(ab_dataset_reduced, Department == "Sales")

# Remove the 'Department' column since it's no longer needed after filtering for Sales
sales_subset$Department <- NULL

# Remove single-level factors which can cause issues in modeling
sales_subset <- sales_subset[, sapply(sales_subset, function(x) !is.factor(x) || length(levels(x)) > 1)]

# Split data into training and testing sets
set.seed(123) # For reproducibility
index <- sample(1:nrow(sales_subset), round(0.75 * nrow(sales_subset)))
train_data <- sales_subset[index, ]
test_data_sales <- sales_subset[-index, ]

# Balancing the training dataset by downsampling the majority class
table_train <- table(train_data$Attrition_numeric)
min_class_size <- min(table_train)

# Subsetting each class to separate the majority and minority classes
train_data_0 <- train_data[train_data$Attrition_numeric == 0, ]
train_data_1 <- train_data[train_data$Attrition_numeric == 1, ]

# Sampling from the majority class to balance with the minority class
train_data_0_downsampled <- train_data_0[sample(1:nrow(train_data_0), min_class_size), ]
train_data_balanced <- rbind(train_data_0_downsampled, train_data_1)

# Shuffle rows to mix classes well, ensuring the model isn't biased towards the original order
set.seed(123)
train_data_balanced <- train_data_balanced[sample(nrow(train_data_balanced)), ]

# Fit logistic regression model
tryCatch({
  sales_model <- glm(Attrition_numeric ~ ., data = train_data_balanced, family = binomial())
  print(summary(sales_model))
}, error = function(e) {
  print("Error fitting model:")
  print(e$message)
})

# Predict and evaluate with confusion matrix
pred_prob <- predict(sales_model, test_data_sales, type = "response")
pred_class <- ifelse(pred_prob > 0.5, 1, 0)
confusion_matrix <- table(Predicted = pred_class, Actual = test_data_sales$Attrition_numeric)
print(confusion_matrix)

# Calculate and print accuracy
accuracy <- sum(diag(confusion_matrix)) / sum(confusion_matrix)
print(paste("Accuracy:", accuracy))


```



```{r}
# Check for NAs or unusual values in the test dataset
summary(test_data_sales)

```





```{r}

# Predict and create the confusion matrix using the caret package
predictions <- predict(sales_model, test_data_sales, type = "response")
predicted_classes <- ifelse(predictions > 0.5, 1, 0)
conf_matrix <- confusionMatrix(factor(predicted_classes), factor(test_data_sales$Attrition_numeric))

# Print the detailed confusion matrix and associated statistics
print(conf_matrix)

```







```{r roc-plot, echo=FALSE, message=FALSE, warning=FALSE}

# Check the number of rows in the test dataset
print(paste("Number of rows in test dataset:", nrow(test_data_sales)))

# Check the length of the prediction vector
print(paste("Length of prediction vector:", length(pred_prob)))

# If they don't match, there's a need to regenerate the predictions
if (length(pred_prob) != nrow(test_data_sales)) {
  print("Mismatch in vector lengths. Regenerating predictions.")
  
  # Ensure the model and test data are correctly aligned
  pred_prob <- predict(sales_model, newdata = test_data_sales, type = "response")
  
  # Recheck lengths
  print(paste("Rechecked length of prediction vector:", length(pred_prob)))
  
  # Continue with ROC and AUC calculation only if lengths match
  if (length(pred_prob) == nrow(test_data_sales)) {
    # Calculate ROC curve
    roc_curve <- roc(response = test_data_sales$Attrition_numeric, predictor = pred_prob)
    
    # Plot ROC curve
    plot(roc_curve, main = "ROC Curve for Sales Department Attrition Model")
    abline(a = 0, b = 1, col = "red", lty = 2)  # Adding a reference line
    
    # Calculate AUC
    auc_value <- auc(roc_curve)
    print(paste("AUC:", auc_value))
  } else {
    print("Error: Prediction vector still does not match the number of rows in test data.")
  }
} else {
  # Calculate ROC curve and AUC if lengths are confirmed to match
  roc_curve <- roc(response = test_data_sales$Attrition_numeric, predictor = pred_prob)
  
  # Plot ROC curve
  plot(roc_curve, main = "ROC Curve for Sales Department Attrition Model")
  abline(a = 0, b = 1, col = "red", lty = 2)  # Adding a reference line
  
  # Calculate AUC
  auc_value <- auc(roc_curve)
  print(paste("AUC:", auc_value))
}



```




# Research and Development subset


```{r}
# Subset data for the Research & Development department from the correct dataset
rd_subset <- subset(ab_dataset_reduced, Department == "Research & Development")

# Remove the 'Department' column as it is not needed after the subset by department
rd_subset$Department <- NULL

# Remove single-level factors which can cause issues in modeling
rd_subset <- rd_subset[, sapply(rd_subset, function(x) !is.factor(x) || length(levels(x)) > 1)]

# Split data into training and testing sets
set.seed(123)  # Set seed for reproducibility
index <- sample(1:nrow(rd_subset), round(0.75 * nrow(rd_subset)))
train_data <- rd_subset[index, ]
test_data_rd <- rd_subset[-index, ]

# Balancing the training dataset by downsampling the majority class
table_train <- table(train_data$Attrition_numeric)
min_class_size <- min(table_train)

# Subsetting each class to separate the majority and minority classes
train_data_0 <- train_data[train_data$Attrition_numeric == 0, ]
train_data_1 <- train_data[train_data$Attrition_numeric == 1, ]

# Sampling from the majority class to balance with the minority class
train_data_0_downsampled <- train_data_0[sample(1:nrow(train_data_0), min_class_size), ]
train_data_balanced <- rbind(train_data_0_downsampled, train_data_1)

# Shuffle rows to mix classes well, ensuring the model isn't biased towards the original order
set.seed(123)
train_data_balanced <- train_data_balanced[sample(nrow(train_data_balanced)), ]

# Fit logistic regression model
tryCatch({
  rd_model <- glm(Attrition_numeric ~ ., data = train_data_balanced, family = binomial())
  print(summary(rd_model))
}, error = function(e) {
  print("Error fitting model:")
  print(e$message)
})

# Predict and evaluate with a confusion matrix
pred_prob <- predict(rd_model, test_data_rd, type = "response")
pred_class <- ifelse(pred_prob > 0.5, 1, 0)
confusion_matrix <- table(Predicted = pred_class, Actual = test_data_rd$Attrition_numeric)
print(confusion_matrix)

# Calculate and print accuracy
accuracy <- sum(diag(confusion_matrix)) / sum(confusion_matrix)
print(paste("Accuracy:", accuracy))

```

# Research and Development confusion matrix


```{r}

# Predict and create the confusion matrix using the caret package
predictions_rd <- predict(rd_model, test_data_rd, type = "response")
predicted_classes_rd <- ifelse(predictions_rd > 0.5, 1, 0)
conf_matrix_rd <- confusionMatrix(factor(predicted_classes_rd), factor(test_data_rd$Attrition_numeric))

# Print the detailed confusion matrix and associated statistics
print(conf_matrix_rd)

```

# R&D subset ROC curve

```{r}

  # Predict probabilities for the test set
  pred_prob <- NULL
  tryCatch({
    pred_prob <- predict(rd_model, newdata = test_data_rd, type = "response")
  }, error = function(e) {
    print("Error in prediction:")
    print(e$message)
  })
  
  # Only proceed if pred_prob has been successfully created
  if (!is.null(pred_prob) && length(pred_prob) == length(test_data_rd$Attrition_numeric)) {
    library(pROC)
    
    # Generate ROC curve
    roc_curve <- roc(response = test_data_rd$Attrition_numeric, predictor = pred_prob)
    
    # Plot ROC curve
    plot(roc_curve, main = "ROC Curve for R&D Logistic Model")
    abline(a = 0, b = 1, col = "red", lty = 2)  # Diagonal line for reference
    
    # Calculate AUC
    auc_value <- auc(roc_curve)
    print(paste("AUC:", auc_value))
    
    # Optional: Add AUC to the plot
    legend("bottomright", legend = paste("AUC =", round(auc_value, 4)), box.lty = 1, col = "black", bty = "n")
  } else {
    print("Error: Unable to generate ROC curve due to previous error or mismatch in data lengths.")
  }

```




```{r}

```





# Human Resources Subset



```{r}
# Subset data for the Human Resources department from the updated dataset name
hr_subset <- subset(ab_dataset_reduced, Department == "Human Resources")

# Remove the 'Department' column as it is no longer needed after the subset by department
hr_subset$Department <- NULL

# Remove single-level factors which can cause issues in modeling
hr_subset <- hr_subset[, sapply(hr_subset, function(x) !is.factor(x) || length(levels(x)) > 1)]

# Split data into training and testing sets
set.seed(123)  # Set seed for reproducibility
index <- sample(1:nrow(hr_subset), round(0.75 * nrow(hr_subset)))
train_data <- hr_subset[index, ]
test_data_hr <- hr_subset[-index, ]

# Balancing the training dataset by downsampling the majority class
table_train <- table(train_data$Attrition_numeric)
min_class_size <- min(table_train)

# Subsetting each class to separate the majority and minority classes
train_data_0 <- train_data[train_data$Attrition_numeric == 0, ]
train_data_1 <- train_data[train_data$Attrition_numeric == 1, ]

# Sampling from the majority class to balance with the minority class
train_data_0_downsampled <- train_data_0[sample(1:nrow(train_data_0), min_class_size), ]
train_data_balanced <- rbind(train_data_0_downsampled, train_data_1)

# Shuffle rows to mix classes well, ensuring the model isn't biased towards the original order
set.seed(123)
train_data_balanced <- train_data_balanced[sample(nrow(train_data_balanced)), ]

# Fit logistic regression model
tryCatch({
  hr_model <- glm(Attrition_numeric ~ ., data = train_data_balanced, family = binomial())
  print(summary(hr_model))
}, error = function(e) {
  print("Error fitting model:")
  print(e$message)
})

# Predict and evaluate with a confusion matrix
pred_prob <- predict(hr_model, test_data_hr, type = "response")
pred_class <- ifelse(pred_prob > 0.5, 1, 0)
confusion_matrix <- table(Predicted = pred_class, Actual = test_data_hr$Attrition_numeric)
print(confusion_matrix)

# Calculate and print accuracy
accuracy <- sum(diag(confusion_matrix)) / sum(confusion_matrix)
print(paste("Accuracy:", accuracy))


```



```{r}

# Predict and create the confusion matrix using the caret package
predictions_hr <- predict(hr_model, test_data_hr, type = "response")
predicted_classes_hr <- ifelse(predictions_hr > 0.5, 1, 0)
conf_matrix_hr <- confusionMatrix(factor(predicted_classes_hr), factor(test_data_hr$Attrition_numeric))

# Print the detailed confusion matrix and associated statistics
print(conf_matrix_hr)

```




```{r}
# Fit logistic regression model
tryCatch({
  hr_model <- glm(Attrition_numeric ~ ., data = train_data_balanced, family = binomial())
}, error = function(e) {
  print("Error in model fitting:")
  print(e$message)
})

# Make predictions on the test dataset
tryCatch({
  pred_prob <- predict(hr_model, newdata = test_data_hr, type = "response")
}, error = function(e) {
  print("Error in prediction:")
  print(e$message)
})

# Re-check the lengths
print(paste("Re-checked length of actual test responses:", length(test_data_hr$Attrition_numeric)))
print(paste("Re-checked length of predicted probabilities:", length(pred_prob)))

```



# HR Subset ROC Curve   



```{r}
# Only proceed if lengths match
if (length(test_data_hr$Attrition_numeric) == length(pred_prob)) {
  library(pROC)

  # Generate ROC curve
  roc_curve <- roc(response = test_data_hr$Attrition_numeric, predictor = pred_prob)

  # Plot ROC curve
  plot(roc_curve, main = "ROC Curve for HR Department Logistic Regression Model")
  abline(a = 0, b = 1, col = "red")  # Diagonal reference line

  # Calculate AUC
  auc_value <- auc(roc_curve)
  print(paste("AUC:", auc_value))
  
  # Optional: Add AUC to the plot
  legend("bottomright", legend = paste("AUC =", round(auc_value, 4)), box.lty = 1, col = "black", bty = "n")
} else {
  print("Length mismatch between test responses and predictions. Unable to generate ROC curve.")
}

```



# Comparison of Model accuracies across subsets (HR, Sales, R&D)


```{r}

# Define a function to summarize and print model results
print_model_summary <- function(department, model, test_data) {
  cat("\n----------------------------------------\n")
  cat("Department:", department, "\n")
  pred_prob <- predict(model, test_data, type = "response")
  pred_class <- ifelse(pred_prob > 0.5, 1, 0)
  confusion_matrix <- table(Predicted = pred_class, Actual = test_data$Attrition_numeric)
  accuracy <- sum(diag(confusion_matrix)) / sum(confusion_matrix)
  
  cat("Confusion Matrix:\n")
  print(confusion_matrix)
  cat("Accuracy:", sprintf("%.2f%%", accuracy * 100), "\n")
}


# Print summaries
print_model_summary("Sales", sales_model, test_data_sales)
print_model_summary("Research & Development", rd_model, test_data_rd)
print_model_summary("Human Resources", hr_model, test_data_hr)

```



# Propensity Score Matching



```{r}

data <- ab_dataset_reduced
# Calculate propensity scores
ps_model <- glm(Attrition_numeric ~ ., data = trainingdata_balanced_ab, family = "binomial")

# Perform matching
match_data <- matchit(Attrition_numeric ~ ., data = trainingdata_balanced_ab, method = "nearest", model = ps_model)

# Extract matched data
matched_data <- match.data(match_data)

set.seed(123)  # For reproducibility
split <- sample.split(matched_data$Attrition_numeric, SplitRatio = 0.7)
train_data <- subset(matched_data, split == TRUE)
test_data <- subset(matched_data, split == FALSE)
# Fit logistic regression on the balanced training data

reduced_model <- glm(Attrition_numeric ~ .,
                     data = trainingdata_balanced_ab, family = binomial())


# Calculate VIF to check for multicollinearity

vif(reduced_model)  # A VIF value greater than 5 is typically cause for concern
# Predict using the reduced model
predicted_probabilities <- predict(reduced_model, newdata = testingdata_reduced_ab, type = "response")
predicted_classes <- ifelse(predicted_probabilities > 0.5, 1, 0)

# Generate a confusion matrix to evaluate the model

confusionMatrix(as.factor(predicted_classes), as.factor(testingdata_reduced_ab$Attrition_numeric))

```


# Variable Interaction for Logistic regression


```{r}

# Set seed for reproducibility
set.seed(123)

# Splitting the dataset into training and testing sets (70:30)
split <- createDataPartition(training_set$Attrition_numeric, p = 0.7, list = FALSE, times = 1)
trainingset_interaction <- training_set[split, ]
testset_interaction <- training_set[-split, ]

# Balancing the 'Attrition_numeric' class in the training set using upSample
balanceddata_interaction <- upSample(x = trainingset_interaction[, names(trainingset_interaction) != "Attrition_numeric"],
                                     y = trainingset_interaction$Attrition_numeric)

# Check if the target variable is in the dataset
# Since upSample does not change the target variable name, no renaming is needed
balanceddata_interaction$Attrition_numeric <- as.factor(balanceddata_interaction$Class)
balanceddata_interaction$Class <- NULL  # Clean up, just in case

# Fit logistic regression model on balanced training set
logistic_model_interaction <- glm(Attrition_numeric ~ Department + EducationField + Age + 
                                  BusinessTravel + DailyRate + DistanceFromHome + Education + 
                                  EnvironmentSatisfaction + Gender + HourlyRate + JobInvolvement + 
                                  JobSatisfaction + MaritalStatus + MonthlyRate + NumCompaniesWorked + 
                                  OverTime + PercentSalaryHike + PerformanceRating + RelationshipSatisfaction + 
                                  StockOptionLevel + TrainingTimesLastYear + WorkLifeBalance + 
                                  YearsSinceLastPromotion + YearsWithCurrManager +
                                  MonthlyIncome * JobLevel + TotalWorkingYears * JobLevel + 
                                  YearsInCurrentRole * YearsAtCompany, 
                                  data = balanceddata_interaction, family = binomial())

# Print out the results of the logistic regression model
summary(logistic_model_interaction)

# Predicting on the test set
predictions <- predict(logistic_model_interaction, newdata = testset_interaction, type = "response")
predicted_classes <- ifelse(predictions > 0.5, 1, 0)

# Evaluate model performance
confusionMatrix <- table(Predicted = predicted_classes, Actual = testset_interaction$Attrition_numeric)
accuracy <- sum(diag(confusionMatrix)) / sum(confusionMatrix)

# Print the confusion matrix and accuracy
print(confusionMatrix)
print(paste("Accuracy:", accuracy))


```



# Interpretation of Coefficients based on the interaction Logistic model - 

OverTimeYes (Coefficient: 0.9370, p-value: < 2e-16)
Most significant predictor of attrition.
Employees working overtime are more likely to leave.
Age (Coefficient: -0.4074, p-value: < 2e-16)
Significant negative predictor.
Older employees are less likely to leave.
BusinessTravelTravel_Frequently (Coefficient: 1.4106, p-value: < 2e-16)
Frequent business travelers are more likely to leave.
DailyRate (Coefficient: -0.1841, p-value: 1.10e-06)
Higher daily rates are associated with lower attrition.
EnvironmentSatisfaction (Coefficient: -0.1985, p-value: 1.80e-06)
Higher satisfaction leads to lower attrition.
BusinessTravelTravel_Rarely (Coefficient: 0.6634, p-value: 3.30e-06)
Less frequent travel still predicts higher attrition compared to no travel.
YearsInCurrentRole (Coefficient: -0.3205, p-value: 1.88e-06)
Longer time in current role associates with lower attrition.
JobInvolvement (Coefficient: -0.1804, p-value: 5.82e-07)
Higher job involvement is linked to lower attrition.
JobSatisfaction (Coefficient: -0.1237, p-value: 0.000820)
Higher job satisfaction reduces attrition.
DistanceFromHome (Coefficient: 0.1253, p-value: 0.002589)
Longer commutes increase the likelihood of leaving.







# Confusion Matrix for Logistic model with Variable interaction



```{r}

# Convert predicted_classes and actual values to factors to ensure compatibility with confusionMatrix function
predicted_factors <- factor(predicted_classes, levels = c(0, 1))
actual_factors <- factor(testset_interaction$Attrition_numeric, levels = c(0, 1))

evaluation_results <- confusionMatrix(predicted_factors, actual_factors)

# Print the evaluation results which include confusion matrix, overall statistics, and class statistics
print(evaluation_results)

```






```{r echo=TRUE, message=FALSE, warning=FALSE}
library(ggplot2)
library(dplyr)

model_summary <- summary(logistic_model_interaction)
# Create a data frame of variables, their coefficients, and p-values
coefficients <- model_summary$coefficients
data <- as.data.frame(coefficients)
data$Variable <- rownames(coefficients)
rownames(data) <- NULL  # Clean up row names

# Add direction based on the sign of the coefficient
data$Direction <- ifelse(data$Estimate > 0, "Positive", "Negative")

# Remove the intercept and split the data frame into positive and negative coefficients
positive_coeffs <- data %>%
  filter(Estimate > 0 & Variable != "(Intercept)") %>%
  arrange(desc(Estimate))

negative_coeffs <- data %>%
  filter(Estimate < 0 & Variable != "(Intercept)") %>%
  arrange(Estimate)

# Define colors for direction
positive_color <- "#EF553B"
negative_color <- "#00CC96"

# Function to clean up the plot look
clean_theme <- function(p) {
  p + theme_minimal() +
    theme(
      text = element_text(size = 16),
      axis.title.y = element_text(size = 18),
      axis.text.y = element_text(size = 12),
      axis.ticks.y = element_blank(),
      panel.grid.major.y = element_blank(),
      panel.grid.minor.y = element_blank(),
      legend.position = "none"
    )
}

# Function to add coefficient labels inside the bars
add_coefficient_labels <- function(p, data) {
  p + geom_text(
    data = data,
    aes(label = sprintf("%.2f", Estimate), x = reorder(Variable, Estimate), y = ifelse(Direction == "Positive", Estimate - 0.05 * max(data$Estimate), Estimate + 0.05 * abs(min(data$Estimate)))),
    size = 3,
    hjust = ifelse(data$Direction == "Positive", 1.1, -0.1),
    color = "white"
  )
}

# Create the positive coefficients bar plot
p1 <- ggplot(positive_coeffs, aes(x = reorder(Variable, Estimate), y = Estimate, fill = Direction)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  scale_fill_manual(values = positive_color) +
  labs(title = "Positive Coefficients for Attrition")

# Create the negative coefficients bar plot
p2 <- ggplot(negative_coeffs, aes(x = reorder(Variable, -Estimate), y = Estimate, fill = Direction)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  scale_fill_manual(values = negative_color) +
  labs(title = "Negative Coefficients for Attrition")

# Apply the custom theme to the plot
p1 <- clean_theme(p1)
p1 <- add_coefficient_labels(p1, positive_coeffs)

# Apply the custom theme to the plot
p2 <- clean_theme(p2)
p2 <- add_coefficient_labels(p2, negative_coeffs)

# Print the plots
print(p1)
print(p2)



# Convert predicted_classes and actual values to factors to ensure compatibility with confusionMatrix function
predicted_factors <- factor(predicted_classes, levels = c(0, 1))
actual_factors <- factor(testset_interaction$Attrition_numeric, levels = c(0, 1))

# Load the caret package if you haven't already
library(caret)

evaluation_results <- confusionMatrix(predicted_factors, actual_factors)

# Print the evaluation results which include confusion matrix, overall statistics, and class statistics
print(evaluation_results)


```







# Decision Tree model with selected Variables (Without Jobrole, Daily rate, monthly rate etc)



```{r}

# Building the decision tree model 
decision_tree_model <- rpart(Attrition_numeric ~ .,
                             data = trainingdata_balanced_ab, 
                             method = "class")

```




# Decision Tree predictions


```{r}

# Predicting using the decision tree model
tree_predictions <- predict(decision_tree_model, newdata = testingdata_reduced_ab, type = "class")

# Convert predictions to factor to ensure compatibility with confusionMatrix function
# It's essential that the levels match those in the actual outcome variable
tree_predictions_factor <- factor(tree_predictions, levels = levels(testingdata_reduced_ab$Attrition_numeric))

# Evaluation using confusion matrix
tree_evaluation_results <- confusionMatrix(tree_predictions_factor, testingdata_reduced_ab$Attrition_numeric)

# Print the evaluation results
print(tree_evaluation_results)

```

# Cross Validation for Decision Trees


```{r}
# Set up cross-validation settings
train_control <- trainControl(
  method = "cv",          # Use k-fold cross-validation
  number = 5,            # Number of folds in k-fold cross-validation
  savePredictions = "final",
  classProbs = TRUE       # Store class probabilities (useful for ROC, AUC)
)

```






```{r}
# Check current levels of the factor
levels(trainingdata_balanced_ab$Attrition_numeric)

# Adjust the levels to be valid R variable names
levels(trainingdata_balanced_ab$Attrition_numeric) <- make.names(levels(trainingdata_balanced_ab$Attrition_numeric))

# Verify the changes
levels(trainingdata_balanced_ab$Attrition_numeric)

```

# 10 Fold Cross validation


```{r}


# Set up cross-validation settings
train_control <- trainControl(
  method = "cv",          # Use k-fold cross-validation
  number = 10,            # Number of folds in k-fold cross-validation
  savePredictions = "final",
  classProbs = TRUE       # Store class probabilities
)

# Train the model with cross-validation
set.seed(123)  # For reproducibility
decision_tree_model_cv <- train(
  Attrition_numeric ~ .,   # Formula
  data = trainingdata_balanced_ab,  # Training data
  method = "rpart",        # Training method (decision tree)
  trControl = train_control,  # Control object
  metric = "Accuracy"      # Performance metric
)

# Print the model summary
print(decision_tree_model_cv)


```


```{r}

# Extract predictions and observed values from the model
predictions <- decision_tree_model_cv$pred

# Select the final predicted classes 
final_predictions <- predictions$pred

# The observed values
observed_values <- predictions$obs

# Generate the confusion matrix
conf_matrix <- confusionMatrix(data = factor(final_predictions, levels = c("X0", "X1")),
                               reference = factor(observed_values, levels = c("X0", "X1")))

# Print the confusion matrix
print(conf_matrix)


```



```{r}

# Extract results
results <- decision_tree_model_cv$results

# Plotting the accuracy across different 'cp' values
ggplot(data = results, aes(x = cp, y = Accuracy)) +
  geom_line() +  # Line plot to show the trend
  geom_point() + # Points to mark each tested 'cp' value
  labs(title = "Accuracy vs Complexity Parameter (cp)",
       x = "Complexity Parameter (cp)",
       y = "Accuracy") +
  theme_minimal()  # Using a minimal theme for cleaner visualization

```



# Decision Tree Cross validated optimal CP value

```{r}
# Train the decision tree model using rpart with the optimal cp value
decision_tree <- rpart(Attrition_numeric ~ ., 
                       data = trainingdata_balanced_ab, 
                       method = "class", 
                       control = rpart.control(cp = 0.01978022))

# If you wish to train using the entire dataset and visualize, otherwise skip this
# and use the model from the train() function for accuracy results.

```



# Decision Tree Cross validated Plot


```{r}
# Plot the decision tree
rpart.plot(decision_tree, 
           main = "Decision Tree Plot", 
           extra = 104,  # Display splits and class distribution
           under = TRUE,  # Display the numbers under the node
           faclen = 0)  # Factor length, use 0 to auto adjust

```



# Decision Tree plot with all variables


```{r}



# Plotting the decision tree
rpart.plot(decision_tree_model)

```



# Decision Tree Variable Importance


```{r}

variable_importance <- decision_tree_model$variable.importance

# Order variables by importance in descending order
ordered_importance <- sort(variable_importance, decreasing = TRUE)

# Print the ordered variable importance
ordered_importance

```


# Variable Importance plot



```{r}

# Provided variable importance scores
variable_importance <- c(
  Age = 785.6839521, OverTime = 533.4903002, StockOptionLevel = 369.9393083, 
  MaritalStatus = 276.8573304, BusinessTravel = 109.6644593, Education = 65.3445040, 
  YearsAtCompany = 64.4883620, PercentSalaryHike = 18.5971663, TrainingTimesLastYear = 16.7410552, 
  NumCompaniesWorked = 15.4888790, EnvironmentSatisfaction = 13.1240286, EducationField = 4.1421277, 
  YearsSinceLastPromotion = 0.3944884
)

# Sort the importance scores in descending order
sorted_importance <- sort(variable_importance, decreasing = TRUE)

# Print sorted variable importance
print(sorted_importance)

# Create a data frame for plotting
importance_df <- data.frame(Variable = names(sorted_importance), Importance = sorted_importance)

# Plot using ggplot2 without reversal
ggplot(importance_df, aes(x = reorder(Variable, Importance), y = Importance, fill = Importance)) +
  geom_col() +  # Use geom_col() which is equivalent to geom_bar(stat="identity")
  scale_fill_gradient(low = "blue", high = "red") +  # Color gradient from blue to red
  coord_flip() +  # Flip coordinates for horizontal bar plot
  xlab("Variable") + 
  ylab("Importance") +
  ggtitle("Variable Importance from Model") +
  theme_minimal() +  # Apply a minimal theme
  theme(legend.position = "none")  # Hide the legend if not required

```



#Full Decision Tree model with all predictors


```{r}


cp_value <- 0.01  # Adjust this as necessary

# Rebuild the decision tree model with the new complexity parameter
tree_model <- rpart(Attrition_numeric ~ ., data = training_set, method = "class", cp = cp_value)

# Plot the decision tree with adjusted parameters
rpart.plot(tree_model, main="Decision Tree for Attrition Prediction",
           box.palette="RdYlGn",       # Use a red-yellow-green color palette
           shadow.col="gray",          # Add shadow for a 3D effect
           cex=0.65,                    # Adjust text size as needed
           tweak= 1,                  # Adjust the overall size of plot elements
           fallen.leaves=FALSE,         # Align leaf nodes at the bottom of the graph
           type=4,                     # Use enhanced text and splits
           extra=104)                  # Display extra node information

# Evaluate the model on the testing set
predictions <- predict(tree_model, newdata = testing_set, type = "class")
confusionMatrix(predictions, testing_set$Attrition_numeric)


```


```{r}
# Extract variable importance from the rpart model
var_importance <- as.data.frame(tree_model$variable.importance)

# Give names to the dataframe
names(var_importance) <- c("Importance")

# Sort the variables by importance
var_importance_sorted <- var_importance[order(-var_importance$Importance), , drop = FALSE]

# Create a barplot
barplot(var_importance_sorted$Importance, main = "Variable Importance",
        xlab = "Variables", ylab = "Importance", las = 2, cex.names = 0.7)

# If you have a lot of variables and the names are cluttered, you can increase the margin size
par(mar = c(12, 5, 4, 2) + 0.1)  # Adjust the bottom margin to fit variable names
barplot(var_importance_sorted$Importance, main = "Variable Importance",
        xlab = "Variables", ylab = "Importance", las = 2, cex.names = 0.7)

```




# Variable importance for DT model 2 (With full predictors)


```{r improved-bar-plot, echo=TRUE, message=FALSE, warning=FALSE}


var_importance_sorted <- data.frame(
  Variable = rownames(var_importance_sorted),
  Importance = var_importance_sorted$Importance
)

ggplot(var_importance_sorted, aes(x = reorder(Variable, Importance), y = Importance)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() + # Flip coordinates to make bars horizontal
  labs(title = "Variable Importance", 
       x = "Importance", 
       y = "") + # Y label is not needed since variables are on the Y axis
  theme_minimal() +
  theme(axis.text.y = element_text(angle = 0, hjust = 1, size = 8), # Adjust the text size as needed
        axis.title.y = element_blank(), # Remove the Y axis title
        plot.title = element_text(hjust = 0.5)) # Center the plot title


```

# Splitting of data into subsets for Sales, Human Resource and Research and development for Decision Trees


```{r}

# Create subsets for each department
sales_data <- subset(ab_dataset_reduced, Department == "Sales")
hr_data <- subset(ab_dataset_reduced, Department == "Human Resources")
rnd_data <- subset(ab_dataset_reduced, Department == "Research & Development")

# Function to process a department dataset
process_department_data <- function(data) {
  set.seed(123) # for reproducibility
  
  # Split data into training and testing
  index <- createDataPartition(y = data$Attrition_numeric, p = 0.7, list = FALSE)
  training_set <- data[index, ]
  testing_set <- data[-index, ]
  
  # Balance attrition in the training set
  downsampled_data <- downSample(x = training_set[, -which(names(training_set) == "Attrition_numeric")],
                                 y = training_set$Attrition_numeric)
  
  # Fix the outcome variable name after downsampling
  downsampled_data$Attrition_numeric <- downsampled_data$Class
  downsampled_data$Class <- NULL  # Remove the 'Class' column as it's no longer needed

  # Convert Attrition_numeric to a factor (if not already)
  downsampled_data$Attrition_numeric <- factor(downsampled_data$Attrition_numeric, levels = c(0, 1))


  # Run a decision tree model
  tree_model <- rpart(Attrition_numeric ~ ., data = downsampled_data, method = "class")

  # Predict on the testing set
  predictions <- predict(tree_model, newdata = testing_set, type = "class")
  
  # Evaluate the model
  confusion <- confusionMatrix(predictions, testing_set$Attrition_numeric)
  
  # Return the model and its evaluation
  return(list(model = tree_model, evaluation = confusion))
}

# Process each department dataset
sales_results <- process_department_data(sales_data)
hr_results <- process_department_data(hr_data)
rnd_results <- process_department_data(rnd_data)

  

```





# Predictions on test set and Confusion Matrix


```{r}

# Process each department dataset
sales_results <- process_department_data(sales_data)
hr_results <- process_department_data(hr_data)
rnd_results <- process_department_data(rnd_data)

# Print the confusion matrix for the Sales department
print("Confusion Matrix for Sales Department:")
print(sales_results$evaluation)

# Print the confusion matrix for the HR department
print("Confusion Matrix for HR Department:")
print(hr_results$evaluation)

# Print the confusion matrix for the R&D department
print("Confusion Matrix for R&D Department:")
print(rnd_results$evaluation)

```


# Variance Importance for Sales, HR and R&D

```{r}
# Calculate variable importance for each department's model
sales_var_imp <- varImp(sales_results$model, scale = FALSE)
hr_var_imp <- varImp(hr_results$model, scale = FALSE)
rnd_var_imp <- varImp(rnd_results$model, scale = FALSE)

# Print the variable importance for Sales department
print("Variable Importance for Sales Department:")
print(sales_var_imp)

# Print the variable importance for HR department
print("Variable Importance for HR Department:")
print(hr_var_imp)

# Print the variable importance for R&D department
print("Variable Importance for R&D Department:")
print(rnd_var_imp)
```





```{r}
# Plot variable importance for Sales department
plot(sales_var_imp, main="Variable Importance - Sales Department")

# Plot variable importance for HR department
plot(hr_var_imp, main="Variable Importance - HR Department")

# Plot variable importance for R&D department
plot(rnd_var_imp, main="Variable Importance - R&D Department")

```






```{r sales-var-imp-plot, echo=TRUE, message=FALSE, warning=FALSE}


# Create a data frame for the Sales department variable importance
sales_data <- data.frame(
  Variable = c("Age", "BusinessTravel", "DailyRate", "DistanceFromHome",
               "Education", "EnvironmentSatisfaction", "HourlyRate",
               "JobInvolvement", "JobLevel", "JobRole"),
  Importance = c(151.418909, 116.449477, 64.651043, 93.528052,
                 7.479869, 15.834588, 61.812332, 62.588595,
                 7.500696, 27.948940)
)

# Plotting
ggplot(sales_data, aes(x = reorder(Variable, -Importance), y = Importance)) +
  geom_col(fill = "steelblue") +
  labs(title = "Variable Importance for Sales Department",
       x = "Variable",
       y = "Importance") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))



```





```{r hr-var-imp-plot, echo=TRUE, message=FALSE, warning=FALSE}


# Create a data frame for the HR department variable importance
hr_data <- data.frame(
  Variable = c("Age", "BusinessTravel", "DailyRate", "DistanceFromHome",
               "Education", "EducationField", "EnvironmentSatisfaction",
               "HourlyRate", "JobInvolvement", "JobRole"),
  Importance = c(48.130178, 22.913872, 31.306332, 60.253362,
                 10.869100, 46.443127, 61.366961, 26.498422,
                 5.205372, 2.909091)
)

# Plotting
ggplot(hr_data, aes(x = reorder(Variable, -Importance), y = Importance)) +
  geom_col(fill = "darkgreen") +
  labs(title = "Variable Importance for HR Department",
       x = "Variable",
       y = "Importance") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```





```{r rnd-var-imp-plot, echo=TRUE, message=FALSE, warning=FALSE}


# Create a data frame for the R&D department variable importance
rnd_data <- data.frame(
  Variable = c("Age", "BusinessTravel", "DailyRate", "DistanceFromHome",
               "Education", "JobInvolvement", "JobSatisfaction",
               "MaritalStatus", "MonthlyIncome", "MonthlyRate"),
  Importance = c(133.977348, 69.933187, 104.876065, 6.121094,
                 16.026592, 61.172651, 56.868178, 95.858599,
                 128.497173, 49.547281)
)

# Plotting
ggplot(rnd_data, aes(x = reorder(Variable, -Importance), y = Importance)) +
  geom_col(fill = "firebrick") +
  labs(title = "Variable Importance for R&D Department",
       x = "Variable",
       y = "Importance") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```





# Final Logistic and Decision Tree models and their Confusion Matrix

```{r}
# Convert predictions and actual data to factors with consistent levels
predicted_classes_factor <- factor(predicted_classes, levels = c(0, 1))
actual_classes_factor <- factor(testset_interaction$Attrition_numeric, levels = c(0, 1))

```



```{r}
# Calculate and print the confusion matrix and related statistics for logistic regression
conf_matrix_logistic <- confusionMatrix(data = predicted_classes_factor, reference = actual_classes_factor)
print(conf_matrix_logistic)

```


```{r}

tree_predictions <- decision_tree_model_cv$pred
predicted_classes_tree <- factor(tree_predictions$pred, levels = c("X0", "X1"))
actual_classes_tree <- factor(tree_predictions$obs, levels = c("X0", "X1"))

# Calculate and print the confusion matrix and related statistics for decision tree
conf_matrix_tree <- confusionMatrix(data = predicted_classes_tree, reference = actual_classes_tree)
print(conf_matrix_tree)

```




```{r}


# Compute the confusion matrices
conf_matrix_logistic <- confusionMatrix(as.factor(predicted_classes), as.factor(testset_interaction$Attrition_numeric))
conf_matrix_tree <- confusionMatrix(factor(decision_tree_model_cv$pred$pred, levels = c("X0", "X1")),
                                    factor(decision_tree_model_cv$pred$obs, levels = c("X0", "X1")))

# Function to extract metrics from confusion matrix
extract_metrics <- function(cm) {
  data.frame(
    Accuracy = cm$overall['Accuracy'],
    Kappa = cm$overall['Kappa'],
    Precision = cm$byClass['Pos Pred Value'],
    Recall = cm$byClass['Sensitivity'],
    F1 = cm$byClass['F1']
  )
}

# Extract metrics
metrics_logistic <- extract_metrics(conf_matrix_logistic)
metrics_tree <- extract_metrics(conf_matrix_tree)

# Combine metrics into one dataframe for plotting
metrics_df <- rbind(metrics_logistic, metrics_tree)
metrics_df$Model <- c("Logistic Regression", "Decision Tree")
metrics_df <- reshape2::melt(metrics_df, id.vars = "Model")

```





```{r}
# Plotting the metrics with additional enhancements
ggplot(metrics_df, aes(x = variable, y = value, fill = Model)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8)) +
  geom_text(aes(label = round(value, 2)), 
            position = position_dodge(width = 0.8), 
            vjust = -0.25, 
            size = 3.5) +
  labs(title = "Comparison of Model Metrics", x = "Metric", y = "Value") +
  theme_minimal(base_size = 11.5) +  # Increasing base size for better readability
  scale_fill_brewer(palette = "Set1") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom") +  # Adjust legend position
  theme(plot.title = element_text(size = 20, face = "bold"),  # Enhance title style
        axis.title = element_text(size = 14))  # Enhance axis titles

```




```{r}
library(pROC)
library(ggplot2)
library(plotly)

# Function to plot ROC curve and calculate AUC
plot_roc_auc <- function(predicted_probs, actuals, model_name) {
  roc_obj <- roc(actuals, predicted_probs)
  auc_value <- auc(roc_obj)
  
  # Create a data frame for plotting
  roc_data <- data.frame(
    FPR = roc_obj$specificities,
    TPR = roc_obj$sensitivities
  )
  
  # Plot ROC curve using ggplot2
  roc_plot <- ggplot(roc_data, aes(x = FPR, y = TPR)) +
    geom_line(color = ifelse(model_name == "Logistic Regression", "blue", "green"), size = 1.5) +
    geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
    ggtitle(paste("ROC Curve -", model_name)) +
    xlab("False Positive Rate (1 - Specificity)") +
    ylab("True Positive Rate (Sensitivity)") +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5))
  
  # Convert ggplot to plotly object for interactivity
  roc_plotly <- ggplotly(roc_plot) %>% layout(legend = list(x = 0.8, y = 0.2))  # Adjust legend position
  
  # Add AUC value to the plot
  auc_text <- paste("AUC =", round(auc_value, 2))
  
  return(list(plot = roc_plotly, auc = auc_text))
}

# Calculate predicted probabilities for logistic regression model (using "logistic_model_interaction")
predicted_probs_logistic <- predict(logistic_model_interaction, newdata = testset_interaction, type = "response")
# Plot ROC curve and calculate AUC for logistic regression model
roc_logistic <- plot_roc_auc(predicted_probs_logistic, testset_interaction$Attrition_numeric, "Logistic Regression")

# Calculate predicted probabilities for decision tree model
predicted_probs_tree <- predict(decision_tree_model_cv, newdata = testset_interaction, type = "prob")[, "X1"]
# Plot ROC curve and calculate AUC for decision tree model
roc_tree <- plot_roc_auc(predicted_probs_tree, testset_interaction$Attrition_numeric, "Decision Tree")

roc_logistic$auc  # Print AUC for logistic regression model
roc_tree$auc  # Print AUC for decision tree model

roc_logistic$plot  # Display ROC plot for logistic regression model
roc_tree$plot  # Display ROC plot for decision tree model

```



